Robust Bayesian Analysis of Causal Inference in Medical Diagnosis and Treatment

### 2024-07-30 JASA-template.tex
Reviewer 1: explain choice of linear model.
Note importance of having no interaction terms in the model to the reviewer, and make comment in paper.

Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\citep{winship99,HECKMAN1985}.
To do so, we rely on $p$ observed quantities, called
\emph{predictors}, denoted by $X\coloneqq$ $[X_1^{\top}, \dots, X_n^{\top}]^{\top}$
where each $X_i\in\mathbb{R}^p$.
Each $X_i$ is treated as a $p$-dimensional row vector,
so $X$ is a $n\times p$ matrix.
Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)^{\top}$ denote the vector of regression
coefficients
related to the predictors, and let $\beta_T$ denote a regression coefficient related to the
treatment decision.
Then we can define a linear model for the outcome
so that
\begin{equation}
	Y_i =  T_i \beta_{T} + X_i\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$,
independent of $T_i$ and $X_i$.
Note that both $T_i$ and $X_i$ are predictors for $Y_i$ in the above model.
However, when we talk about predictors in this paper, we usually mean just the components of $X_i$.

To decide whether or not to treat a new individual with given predictors,
we are mainly interested in the effect of the treatment on the outcome.
More precisely, the causal effect of a new individual, indexed as $n+1$,
whose outcome $Y_{n+1}$ is not yet observed, and with observed predictors $X_{n+1}=x_{n+1}$, is defined by:
\begin{align}
  &\delta(x_{n+1})\nonumber\\
  &\coloneqq\mathbb{E}(Y_{n+1}\mid X_{n+1}=x_{n+1},T_{n+1} =1) - \mathbb{E}(Y_{n+1}\mid X_{n+1}=x_{n+1},T_{n+1}=0)
\end{align}
For our model, due to linearity of expectation, we have that
\begin{align}
  \delta(x_{n+1})
  =\beta_T+&x_{n+1}\beta+\mathbb{E}(\epsilon_{n+1}\mid X_{n+1}=x_{n+1},T_{n+1} =1) \nonumber\\
  - &x_{n+1}\beta-\mathbb{E}(\epsilon_{n+1}\mid X_{n+1}=x_{n+1},T_{n+1} =0)
  \end{align}
  {and because $\epsilon_{n+1}$ is independent from $X_{n+1}$ and $T_{n+1}$,}
  \begin{align}
  \delta(x_{n+1})
  &=\beta_{T}.
\end{align}
Note that, for this model, the causal effect $\delta(x_{n+1})$
does not depend on the observed value $x_{n+1}$ of $X_{n+1}$.
So, to find the causal effect, we simply need to estimate $\beta_T$.

>>>>

Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\citep{winship99,HECKMAN1985}.
To do so, we rely on $p$ observed quantities, called
\emph{predictors}, denoted by $X\coloneqq$ $[X_1^{\top}, \dots, X_n^{\top}]^{\top}$
where each $X_i\in\mathbb{R}^p$.
Each $X_i$ is treated as a $p$-dimensional row vector,
so $X$ is a $n\times p$ matrix.
Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)^{\top}$ denote the vector of regression
coefficients
related to the predictors, and let $\beta_T$ denote a regression coefficient related to the
treatment decision.
Following the usual approach in the literature
(see for instance \citep{winship99,HECKMAN1985}),
we model the outcome using a linear model
\begin{equation}
	Y_i =  T_i \beta_{T} + X_i\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$,
independent of $T_i$ and $X_i$.
Note that both $T_i$ and $X_i$ are predictors for $Y_i$ in the above model.
However, when we talk about predictors in this paper, we usually mean just the components of $X_i$.

To decide whether or not to treat a new individual with given predictors,
we are mainly interested in the effect of the treatment on the outcome.
More precisely, the causal effect of a new individual, indexed as $n+1$,
whose outcome $Y_{n+1}$ is not yet observed, and with observed predictors $X_{n+1}=x_{n+1}$, is defined by:
\begin{align}
  &\delta(x_{n+1})\nonumber\\
  &\coloneqq\mathbb{E}(Y_{n+1}\mid X_{n+1}=x_{n+1},T_{n+1} =1) - \mathbb{E}(Y_{n+1}\mid X_{n+1}=x_{n+1},T_{n+1}=0)
\end{align}
For our model, due to linearity of expectation, we have that
\begin{align}
  \delta(x_{n+1})
  =\beta_T+&x_{n+1}\beta+\mathbb{E}(\epsilon_{n+1}\mid X_{n+1}=x_{n+1},T_{n+1} =1) \nonumber\\
  - &x_{n+1}\beta-\mathbb{E}(\epsilon_{n+1}\mid X_{n+1}=x_{n+1},T_{n+1} =0)
  \end{align}
  {and because $\epsilon_{n+1}$ is independent from $X_{n+1}$ and $T_{n+1}$,}
  \begin{align}
  \delta(x_{n+1})
  &=\beta_{T}.
\end{align}
Note that, for this model, the causal effect $\delta(x_{n+1})$
does not depend on the observed value $x_{n+1}$ of $X_{n+1}$.
So, to find the causal effect, we simply need to estimate $\beta_T$.
Note that, if interaction terms between $X$ and $T$ were present in
the model (for example a term of the form, say, $\eta XT$ for some
parameter vector $\eta$), that would result in a dependence of the
causal effect on $x_{n+1}$.
